---
title: "Pokémon Dataset Analysis"
subtitle: "Proposal"
author: 
  - name: "David Kennedy (Pezheadpunk)"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Asking and answering two questions about the Pokémon Dataset with data visualizations, and presentation of the findings"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

## Dataset Description

### Provenance 
The data in this dataset came from https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-04-01 and was originally sourced and processed using [PokéAPI](https://pokeapi.co), an open-source, programmatic database containing comprehensive metadata about the Pokémon video game franchise. The dataset combines official in-game mechanical statistics (such as combat attributes and breeding groups) with multimedia design elements pulled directly from community-run encyclopedia archives like [Bulbagarden](https://bulbagarden.net).

### Dimensions and Scale
*   **Observations (Rows):** The dataset catalogs entries tracking unique Pokémon species, distinct forms, and regional variations across **7 historical game generations**.
*   **Variables (Columns):** The schema consists of **22 distinct variables** that map both numerical attributes and structural metadata for each entry.

### Structure & Variable Groupings
The attributes within this file are split across three primary structural categories:

1. **Core Identity Metadata:** Unique tracking metrics including systemic indexes (`id`, `species_id`), common naming conventions (`pokemon`), and generational release timelines (`generation_id`).
2. **Combat & Biological Mechanics:** Core battle metrics mapping scale properties (`height`, `weight`), growth baselines (`base_experience`), combat attributes (`hp`, `attack`, `defense`, `special_attack`, `special_defense`, `speed`), elemental profiles (`type_1`, `type_2`), and biological mechanics (`egg_group_1`, `egg_group_2`).
3. **Aesthetic & Multimedia Assets:** Design-oriented properties mapping primary interface layouts including hexadecimal theme color schemes (`color_1`, `color_2`, `color_f`) alongside direct asset links (`url_icon`, `url_image`).

### Why I Chose This Dataset
I chose this dataset because I have played every generation of Pokemon games since the original on Gameboy. I though it would be interesting to explore a dataset involving data from a game series that I am very familar with as fan.

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
