---
title: "Pokemon Dataset Analysis"
subtitle: "Proposal"
author: 
  - name: "David Kennedy (pezheadpunk)"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Asking and answering two questions about the Pokemon Dataset with data visualizations, and a presentation of the findings"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

pokemon_df.csv

```{r}
#| label: load-dataset
#| message: false
pokemon_data <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2025/2025-04-01/pokemon_df.csv") 
```

## Dataset Description

### Provenance

The data in this dataset came from [tidytuesday](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-04-0), and was originally sourced and processed using [PokéAPI](https://pokeapi.co), an open-source, programmatic database containing comprehensive metadata about the Pokemon video game franchise. The dataset combines official in-game mechanical statistics (such as combat attributes and breeding groups) with multimedia design elements pulled directly from community-run encyclopedia archives like [Bulbagarden](https://bulbagarden.net).

### Dimensions and Structure

```{r}
glimpse(pokemon_data)
```

### Why I Chose This Dataset

I chose this dataset because I have played every generation of Pokemon games since the original on Gameboy. I though it would be interesting to explore a dataset involving data from a game series that I am very familiar with as fan.

## Questions

The two questions we want to answer.

### Question 1

- How does a Pokemon's primary type and generation introduction interact to influence its physical durability compared to its special durability?

### Question 2

- Does a Pokemon's Body Mass Index (BMI) serve as a reliable predictor for its combat speed and physical attack power and follows the biological rule that heavier, bulkier creatures sacrifice speed for raw physical power?

## Analysis plan

### Question 1

Variables Involved:

- type_1 (primary elemental type)

- generation_id (generation introduced)

- hp (base hit points)

- defense (physical defense)

- special_defense

Variable to Be Created:

- Physical Durability = hp + defense

- Special Durability = hp + special_defense

Plan:

To answer the question we will combine the base stats of hp, defense, and special_defense into overall physical and special durability scores for each Pokemon. Then group the data by type and generation to create a multi-panel line graph that tracks how defensive playstyles evolved over the video game generations. If the graph is too busy with all the types, a group of 5 types will be chosen.

### Question 2

Variables Involved:

- weight

- height

- speed

- attack

Variable to Be Created:

- Body Mass Index (BMI) = weight / (height ^ 2)

Plan:

To answer this question we will convert each Pokemon's height and weight into a single Body Mass Index (BMI) metric. Then create two scatter plots that uses point positions, sizes, and colors to show how a Pokemon's body size impacts its speed, and another its physical power. Possibly add in the primary type of the pokemon in some way to also observe which are the "heavier" types.

### Weekly Project Plan Table
| Task         | Status   | Due      | Priority | Summary                        |
| ------------ | -------- | -------- | -------- | ------------------------------ |
| Peer Review  | Complete | 6/17/26  | Low      | Close issues                   |
| Write-Up     | To Do    | 6/26/26  | High     | Write up project               |
| Slides       | To Do    | 7/1/26   | High     | Create slides for presentation |
| Presentation | To Do    | 7/3/26   | High     | Record and post presentation   |
